Assessing Digital Elevation Models with different image overlap using photogrammetry

### 2021-04-29 Paper.tex
Updates from Overleaf

\documentclass{isprs} % isprs class modified 23-04-2019 (Dennis Wittich)
\usepackage{subfigure}
\usepackage{setspace}
\usepackage{geometry} % added 27-02-2014 Markus Englich
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[colorlinks, allcolors=blue]{hyperref}
%\bibliographystyle{apalike}
\usepackage[labelsep=period]{caption}  % added 14-04-2016 Markus Englich - Recommendation by Sebastian Brocks
\usepackage[british]{babel} 
\usepackage[hang]{footmisc}
\def\footnotemargin{1em} % added 08-01-2020 Dennis Wittich
\usepackage{fancyref}
\pagestyle{plain}

@@

First, the complete data set has been aligned on medium quality.
On medium alignment quality the original images are scaled to have 1/2 of the original size.
Next, the dense cloud was built on medium quality (no depth filtering), resulting in a dense cloud of 3,268,264 points.
With medium settings the original photo is scaled to 1/4 before used in the dense cloud generation process.
A mesh was built using the the triangulated irregular network (TIN) algorithm by \citet{axelsson1999processing}. 
finally after the mesh, the Reduce Overlap function was used to reduce the amount images used in the process of photogrammetry. 
Because the final product of the photogrammetry process in this paper is a DEM a 2.5D height field was used when the mesh was generated
The 2.5D height field mesh contains only one Z value for any X, Y coordinate combination, this is also the case in the final DEM.

\textbf{Reduce Overlap.} 
The algorithm of the Reduce Overlap function tries to select a minimal amount of images such that each point of the model is observed from N locations/angles.
If a point is observed multiple times from the same location/angle, than this still counts as one. 
As in this research a height field was used, the algorithm will probably take the vegetation canopy into account in the reconstruction, but lower branches may be lost in the process.
This will has some influence on the Reduce Overlap setting; it becomes more flexible because single height values are for each X, Y coordinate combination is used.
The algorithm and parameters behind the Agisoft Reduce Overlap function are proprietary software and therefore unfortunately unknown.
In Agisoft Metashape the Reduce Overlap had three different settings, low, medium and high. 
All of the settings are used and the remaining images are used for further processing.
In table \ref{tab:ImageProcessing} the result of each Reduce Overlap settings can be found.
The images available in table \ref{tab:ImageProcessing} will be treated as a unique dataset and will be used in the image processing pipeline from Agisoft.

\textbf{Camera Locations.}
After all the new datasets with the images have been aligned the camera locations are estimated, in figure \ref{fig:cameralocation} camera locations can be found. 
The images are matched based on an adapted version of the SIFT algorithm \citep{lowe1999object, AgisoftMetashape}.
All the datasets are processed on medium settings for the image alignment.
None of the images used in the alignment process failed to align.

@@

In table \ref{tab:ImageProcessing} an overview of the image processing results can be found.

\begin{table}[h]
    \centering
    \caption{Image Processing Results}
    \begin{tabular}{@{}cccc@{}}
    \toprule
    \textbf{Setting} & \textbf{Images} & \multicolumn{1}{l}{\textbf{Sparse Cloud}} & \multicolumn{1}{l}{\textbf{Dense Cloud}} \\ \midrule
    -      & 111 & 80,901 & 3,269,264 \\
    Low    & 44  & 37,886 & 4,172,783 \\
    Medium & 70  & 54,556 & 3,313,399 \\
    High   & 75  & 49,692 & 3,276,832 \\ \bottomrule
    % Add processing time!
    \label{tab:ImageProcessing}
\end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{Processing time in seconds for the difference steps in the process.}
    \label{tab:processingTime}
    \begin{tabular}{@{}cccc@{}}
    \toprule
    \textbf{Setting} & \textbf{Sparse Cloud} & \multicolumn{1}{l}{\textbf{Dense Cloud}} & \multicolumn{1}{l}{\textbf{DEM}} \\ 
    \midrule
    -      & 150 & 292 & 17 \\
    Low    & 59  & 443 & 16 \\
    Medium & 83  & 147 & 17 \\
    High   & 129 & 159 & 17 \\ \bottomrule
\end{tabular}
\end{table}

\textbf{DEM.}
The source data for the DEM's are the dense cloud results. 
The dense clouds contain about the same amount of points with the exception for the lowest Reduce Overlap setting, This setting has resulted in the dense cloud with the most points.
The TIN algorithm is used to built the DEM's \citep{axelsson1999processing}.
Because the dense clouds are unique the TIN algorithm also results in four unique DEM's.
These DEM's all have a slightly different extent and resolution.
For further comparison the DEM's have all been cropped on the largest common extent and the pixel resolution was set to 17.5 cm per pixel (1670 x 1290 pixels).
The DEM's will also be scaled to check the differences between the DEM's without looking at the absolute values.
The scaling will be done by subtracting the mean and dividing by the standard deviation.
For further comparison all of the DEM's will be subtracted from the full model DEM.
This will result in three new DEM's that contain differences instead of height values.
These data will be used to inspect the differences between the scaled DEM of the full image dataset and the scaled DEM's of the other datasets.

\section{Results}
In figure \ref{fig:DemPlot_unscaled} the DEM's are visualized using the same color range. 
In this figure the DEM's do no look similar at all.

\begin{figure}[h]
    \centering
    \includegraphics[width=8cm]{DemPlots.png}
    \caption{DEM Results of the Reduce Overlap}
    \label{fig:DemPlot_unscaled}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=8cm]{DemBoxplot.png}
    \caption{Box plots of the DEM Results of the Reduce Overlap}
    \label{fig:BoxPlot_unscaled}
\end{figure}

Figure \ref{fig:DemPlot_scaled} contains the same visualization of DEM plots, only for all the DEM's the data has been scaled.
There is almost no visual difference visible in the figure, the DEM's look quite similar.

@@

The same applies for figure \ref{fig:BoxPlot_scaled}, only now the boxplot contain the scaled data values.
There difference seem to be minimal in the scaled boxplots.

@@

Figure \ref{fig:ViolinPlot} contains three violin plots. For each of the plots the scaled DEM's from the reduces image datasets has been subtracted from the scaled full dataset DEM.
It is visible that the overall error of the scaled DEM's varies from around -5.0 to 5.0.
For the high settings, the differences are more centered around zero, while this is not the case for low and medium settings.
For medium setting the difference is slightly more centered around 0.0 than low settings.

>>>>

\documentclass{isprs} % isprs class modified 23-04-2019 (Dennis Wittich)
\usepackage{subfigure}
\usepackage{setspace}
\usepackage{geometry} % added 27-02-2014 Markus Englich
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[colorlinks, allcolors=blue]{hyperref}
%\bibliographystyle{apalike}
\usepackage[labelsep=period]{caption}  % added 14-04-2016 Markus Englich - Recommendation by Sebastian Brocks
\usepackage[british]{babel} 
\usepackage[hang]{footmisc}
%\setlength{\tabcolsep}{2pt}
\def\footnotemargin{1em} % added 08-01-2020 Dennis Wittich
\usepackage{fancyref}
\pagestyle{plain}

@@

First, the complete data set has been aligned on medium quality.
On medium alignment quality the original images are scaled to have 1/2 of the original size.
%ADD MANUAL
Next, the dense cloud was built on medium quality (no depth filtering), resulting in a dense cloud of 3,268,264 points.
With medium settings the original photo is scaled to 1/4 before used in the dense cloud generation process.
%ADD MANUAL
A mesh was built using the the triangulated irregular network (TIN) algorithm by \citet{axelsson1999processing}. 
finally after the mesh, the Reduce Overlap function was used to reduce the amount images used in the process of photogrammetry. 
Because the final product of the photogrammetry process in this paper is a DEM a 2.5D height field was used when the mesh was generated
The 2.5D height field mesh contains only one Z value for any X, Y coordinate combination, this is also the case in the final DEM.

\textbf{Reduce Overlap.} 
The algorithm of the Reduce Overlap function tries to select a minimal amount of images such that each point of the model is observed from N locations/angles.
If a point is observed multiple times from the same location/angle, than this still counts as one. 
As in this research a height field was used, the algorithm will probably take the vegetation canopy into account in the reconstruction, but lower branches may be lost in the process.
This will have some influence on the Reduce Overlap setting; it becomes more flexible because single height values are used for each X, Y coordinate combination.
The algorithm and parameters behind the Agisoft Reduce Overlap function are proprietary software and therefore unfortunately unknown.
In Agisoft Metashape the Reduce Overlap had three different settings, low, medium and high. 
All of the settings are used and the remaining images are used for further processing.
The images are matched based on an adapted version of the SIFT algorithm \citep{lowe1999object, AgisoftMetashape}.
All the datasets are processed on medium settings for the image alignment.

\textbf{DEM.}
The source data for the DEM's are the dense cloud results. 
The TIN algorithm is used to built the DEM's \citep{axelsson1999processing}.
Because the dense clouds are unique the TIN algorithm also results in four unique DEM's.
These DEM's all have a slightly different extent and resolution.
For further comparison the DEM's have all been cropped on the largest common extent and the pixel resolution was set to 17.5 cm per pixel (1670 x 1290 pixels).
The DEM's were also be scaled to check the differences between the DEM's without looking at the absolute values.
The scaling was be done by subtracting the mean and dividing by the standard deviation.
For further comparison all of the DEM's were be subtracted from the full model DEM.
%%% CHECK THIS
These data will be used to inspect the differences between the scaled DEM of the full image dataset compared to the other DEM's.

\section{Results}
In table \ref{tab:ImageProcessing} the result of image processing for each Reduce Overlap settings can be found. 
The dense clouds contain about the same amount of points with the exception for the lowest Reduce Overlap setting, This setting has resulted in the dense cloud with the most points.

\begin{table}[h]
    \centering
    \caption{Image Processing Results}
    \begin{tabular}{@{}cccc@{}}
    \toprule
    \textbf{\begin{tabular}[c]{@{}c@{}}Reduce Overlap \\ Setting\end{tabular}} &
    \textbf{\begin{tabular}[c]{@{}c@{}}Image\\ Count\end{tabular}} &
    \textbf{Sparse  Cloud} &
    \textbf{Dense Cloud} \\ \midrule
    -      & 111 & 80,901 & 3,269,264 \\
    Low    & 44  & 37,886 & 4,172,783 \\
    Medium & 70  & 54,556 & 3,313,399 \\
    High   & 75  & 49,692 & 3,276,832 \\ \bottomrule
    % Add processing time!
    \label{tab:ImageProcessing}
\end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{Processing time in seconds for the difference steps in the process.}
    \label{tab:processingTime}
    \begin{tabular}{@{}cccc@{}}
    \toprule
    \textbf{Image Count} & \textbf{Sparse  Cloud} & \textbf{Dense Cloud} & \textbf{Dem} \\ \midrule
    111                  & 150                    & 292                  & 17           \\
    44                   & 59                     & 443                  & 16           \\
    70                   & 83                     & 147                  & 17           \\
    75                   & 129                    & 159                  & 17           \\ \bottomrule
    \end{tabular}
\end{table}

After all the new datasets with the images have been aligned the camera locations are estimated, in figure \ref{fig:cameralocation} camera locations can be found. 
None of the images used in the alignment process failed to align.

@@

In figure \ref{fig:DemPlot_unscaled} the DEM's are visualized using the same color range. 
In this figure the DEM's do no look similar at all.

\begin{figure}[h]
    \centering
    \includegraphics[width=8.5cm]{DemPlots.png}
    \caption{DEM Results of the Reduce Overlap}
    \label{fig:DemPlot_unscaled}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=8cm]{DemBoxplot.png}
    \caption{Box plots of the DEM Results of the Reduce Overlap}
    \label{fig:BoxPlot_unscaled}
\end{figure}

Figure \ref{fig:DemPlot_scaled} contains the same visualization of DEM plots, only for all the DEM's the data has been scaled.
There is almost no visual difference visible in the figure, the DEM's look quite similar.
This is confirmed in figure \ref{fig:BoxPlot_scaled}, only now the boxplot contain the scaled data values.
These difference seem to be minimal in the scaled boxplots.

@@

@@

Figure \ref{fig:ViolinPlot} contains three violin plots. For each of the plots the scaled DEM's from the reduced image datasets has been subtracted from the scaled full dataset DEM.
It is visible that the overall scaled error difference of the scaled DEM's varies from around -5.0 to 5.0.
For the high settings, the differences are more centered around zero, while this is not the case for low and medium settings.
For medium setting the difference is slightly more centered around 0.0 than low settings.
